---
title: "BST 260 Full Project"
output: html_document
runtime: shiny
---

This project has two primary objetives: 1) to predict COVID-19 mortality on the county-level through 10/31/2021 using modified versions of previous models by Wu et al, and 2) explore how widespread availabilty of vaccines potentially changed the importance of predictors in our model. Code was modified based on open source provided by: https://github.com/wxwx1993/PM_COVID . This work led to the publication of the following study: "Wu, X., Nethery, R.C., Sabath, M.B., Braun, D. and Dominici, F., 2020. Air pollution and COVID-19 mortality in the United States: strengths and limitations of an ecological regression analysis. Science advances, 6(45), p.eabd4049" We thank all of the authors for making their data public and for enabling our project to be possible.


```{r}
library(dplyr)
library(stringr)
library(RCurl)
library(httr)

date_of_study <- "10-31-2021"

#Use Apr 15 as the Vax cutoff point
date_of_vax <- "04-15-2021"

```

## Import and Prepare Datasets

To isolate all of the necessary data for this project, we used Web Scraping to download county-level COVID mortality data from the Wu et al. study team Github. Our first step was to downloaod the the data used by Wu et al. (covid_histortical). We next downloaded data through the end our of study period 10/31/21, as well as long-term county level PM2.5 and temperature exposure data. 

```{r}
# Historical data
covid_historical <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-30-2020.csv"))
covid_us_historical <- subset(covid_historical, Country_Region == "US" & is.na(FIPS) == F)

```

```{r}
# Import COVID Mortality data from JHU CSSE
df <- read.csv(text = getURL(paste0("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/", date_of_study, ".csv")))
covid <- subset(df, Country_Region == "US")[, 1:12]
covid <- rbind(covid, subset(covid_us_historical, (!(FIPS %in% covid$FIPS)) & Confirmed == 0 & Deaths == 0 & is.na(FIPS) == F))
covid$FIPS <- str_pad(covid$FIPS, 5, pad = "0")

# Import Long-term Exposure PM2.5 and temperature data
county_pm <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_pm25.csv"))
county_temp <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/temp_seasonal_county.csv"))

```



```{r}
# Import census, brfss data as potential confounders
county_census <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/census_county_interpolated.csv"))

GET("https://www.countyhealthrankings.org/sites/default/files/media/document/analytic_data2020.csv", write_disk("county_brfss.csv", overwrite = TRUE))
county_brfss <- read.csv("county_brfss.csv", skip = 1)
county_brfss <- county_brfss[, c('fipscode', 'v011_rawvalue', 'v009_rawvalue')]
names(county_brfss) <- c('fips', 'obese', 'smoke')
county_brfss$fips <- str_pad(county_brfss$fips, 5, pad = "0")

state_test <- read.csv(text = getURL("https://api.covidtracking.com/v1/states/daily.csv"))
state_test <- subset(state_test, date == paste0(substring(str_remove_all(date_of_study, "-"), 5, 8),substring(str_remove_all(date_of_study, "-"), 1, 4)))[, - 38]
statecode <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/statecode.csv"))

hospitals <- read.csv(text = getURL("https://opendata.arcgis.com/datasets/6ac5e325468c4cb9b905f1728d6fbf0f_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D"))
hospitals$BEDS[hospitals$BEDS < 0] <- NA

county_base_mortality <- read.table(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_base_mortality.txt"), sep = "", header = TRUE)
county_old_mortality <- read.table(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_old_mortality.txt"), sep = "", header = TRUE)
county_014_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_014_mortality.txt", sep = "", header = TRUE)
county_1544_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_1544_mortality.txt", sep = "", header = TRUE)
county_4564_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_4564_mortality.txt", sep = "", header = TRUE)

colnames(county_old_mortality)[4] <- c("older_Population")
colnames(county_014_mortality)[4] <- c("014_Population")
colnames(county_1544_mortality)[4] <- c("1544_Population")
colnames(county_4564_mortality)[4] <- c("4564_Population")

county_base_mortality <- merge(county_base_mortality,county_old_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_014_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_1544_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_4564_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)

county_base_mortality$older_pecent <- county_base_mortality$older_Population / county_base_mortality$Population
county_base_mortality$"young_pecent" <- county_base_mortality$"014_Population" / county_base_mortality$Population
county_base_mortality$"prime_pecent" <- county_base_mortality$"1544_Population" / county_base_mortality$Population
county_base_mortality$"mid_pecent" <- county_base_mortality$"4564_Population" / county_base_mortality$Population
county_base_mortality$"older_pecent"[is.na(county_base_mortality$"older_pecent")] <- 0
county_base_mortality$"prime_pecent"[is.na(county_base_mortality$"prime_pecent")] <- 0
county_base_mortality$"mid_pecent"[is.na(county_base_mortality$"mid_pecent")] <- 0
county_base_mortality$"young_pecent"[is.na(county_base_mortality$"young_pecent")] <- 0

# Import NCHS Urban-Rural Classification Scheme for Counties
NCHSURCodes2013 <- read.csv("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/NCHSURCodes2013.csv")
NCHSURCodes2013$FIPS <- str_pad(NCHSURCodes2013$FIPS, 5, pad = "0")

```

## Variable Construction and Dataset Merge & Join

```{r}
# pm2.5 average over 17 years
county_pm_aggregated <- county_pm %>% 
    group_by(fips) %>% 
    summarise(mean_pm25 = mean(pm25))

# temperature and relative humidity average over 17 years
county_temp_aggregated <- county_temp %>% 
  group_by(fips) %>% 
  summarise(mean_winter_temp = mean(winter_tmmx),
            mean_summer_temp = mean(summer_tmmx),
            mean_winter_rm = mean(winter_rmax),
            mean_summer_rm = mean(summer_rmax))

county_pm_aggregated <- merge(county_pm_aggregated,
                              county_temp_aggregated,
                              by = "fips",
                              all.x = TRUE)

county_hospitals_aggregated <- hospitals %>% group_by(COUNTYFIPS) %>% summarise(beds = sum(BEDS, na.rm = TRUE))
county_hospitals_aggregated$COUNTYFIPS <- str_pad(county_hospitals_aggregated$COUNTYFIPS, 5, pad = "0")

county_census_aggregated2 <- subset(county_census, year == 2016)

# merge dataset
county_census_aggregated2$fips <- str_pad(county_census_aggregated2$fips, 5, pad = "0")
county_census_aggregated2 <- merge(county_census_aggregated2,county_brfss,
                                   by = "fips",
                                   all.x = TRUE)

county_pm_aggregated$fips <- str_pad(county_pm_aggregated$fips, 5, pad = "0")
aggregate_pm <- merge(county_pm_aggregated,covid,
                      by.x = "fips",
                      by.y = "FIPS")

aggregate_pm_census <- merge(aggregate_pm,
                             county_census_aggregated2,
                             by.x = "fips",
                             by.y = "fips")

county_base_mortality$County.Code <- str_pad(county_base_mortality$County.Code, 5, pad = "0")
aggregate_pm_census_cdc <- merge(aggregate_pm_census,
                                 county_base_mortality[, c("County.Code", 
                                                           "Population",
                                                           "older_pecent",
                                                           "young_pecent",
                                                           "prime_pecent",
                                                           "mid_pecent")],
                                 by.x = "fips",
                                 by.y = "County.Code",
                                 all.x = TRUE)

aggregate_pm_census_cdc <- aggregate_pm_census_cdc[is.na(aggregate_pm_census_cdc$fips) == F, ]

#Export Data
#write.csv(aggregate_pm_census_cdc,"aggregate_pm_census_cdc_211031.csv", row.names = FALSE)
#write.csv(aggregate_pm_census_cdc,"aggregate_pm_census_cdc_210415.csv", row.names = FALSE)

```

# Part 0: Maps of the US Covariates

```{r}
library(maps)
library(stringr)
library(RColorBrewer)
library(shiny)
library(shinythemes)
library(ggplot2)
library(magrittr)


AllCounty <- map_data("county")


AllCounty$region <- str_to_title(AllCounty$region)
AllCounty$subregion <- str_to_title(AllCounty$subregion)

AllCounty <- left_join (AllCounty, aggregate_pm_census_cdc, by = c("region" = "Province_State", "subregion" = "Admin2")) 


AllCounty %<>% mutate(Mortality = Deaths/Population * 100000,
                     mean_winter_temp = (mean_winter_temp - 273.15)*1.8 + 32,
                   mean_summer_temp = (mean_summer_temp - 273.15)*1.8 + 32,  )


AllCounty %<>% rename(PM2.5 = mean_pm25,
                      Winter_Temperature = mean_winter_temp,
                      Summer_Temperature = mean_summer_temp,
                      Smoking = smoke, 
                      Obesity= obese, 
                      Poverty = poverty, 
                      With_HS_Degrees = no_grad, 
                      Homeowner = owner_occupied, 
                      Black = blk_pct, 
                      Hispanic = hispanic_pct, 
                      Older_Adults = age_pct_65_plus, 
                      Population_Density =population_density,
                      Income =median_household_income, 
                      House_Value = median_house_value
)

```




```{r}

shinyApp(
ui = fluidPage( 
  theme=shinytheme("darkly"),
    titlePanel("A Description of COVID-19 Mortality by County"),
   tabPanel("Range of County-Level Mortality",
        sidebarPanel( p("The map to the right shades the counties in the contiguous US by number of reported COVID-19 deaths per 100,000 inidviduals by October 31, 2021. A darker shade of blue reflects counties with more deaths per capita."),
                 
                 # Add some space between the text above and animated
                 # slider bar below
                 br(),
                 
                 # Input: year slider with basic animation
                 sliderInput("mortality", "Mortality:",
                             min = 0, max = 48,
                             value = 1, 
                             step = 1,
                             sep = "",       # keep years in year format and not 1,960 format
                             ticks = FALSE,  # don't show tick marks on slider bar
                             animate = FALSE)

        
        ),
          mainPanel(
            # Plot
            plotOutput("slider")
         #end sidepane,
        ) #end tab 2
        
    ),
    tabsetPanel(
        tabPanel("Covariates",
                   fluidRow(
    column(4, p("The maps to the right show the spatial distribution of several relevant county-level covarites. The top map depitcts the number of COVID-19 deaths per 100,000 by October 31,2021-- our outcome of interest. The bottom map depicts the distribution of the county-level covariate of interest. These include PM2.5 (our exposure of interest), as well as mean summer and winter temperatures (measured in degrees Fahrenheit), propotion of smokers, proportion with obesity, proportion living in poverty, proportion with at least a high school degree, proportion who own their home, proportion Black, proportion Hispanic, proportion above 65 years of age, average population density, average income, and average home value. To better show the range of values, the population density, income, and average home value were log transformed. Despite the complications introduced on the ease of understanding the numeric value, the shade of the county still follows intuitive logic. That is, as the shade darkens, the covariate increaes."), 
           
    ),
    column(8, plotOutput("plot_static"))
  ),
  fluidRow(
     column(4,
             radioButtons(inputId = "covariate", label = "Select a Covariate",
                         choices = c('PM2.5','Winter_Temperature','Summer_Temperature',
                      'Smoking' , 'Obesity' , 'Poverty' , 'With_HS_Degrees', 'Homeowner', 
                      'Black' , 'Hispanic', "Older_Adults" , 'Population_Density' ,
                      'Income',"House_Value")
                   ),
    ),
    column(8, plotOutput("plot")),
  )
        )
  #end tabset Panel
)), 

server = function(input, output){
  selected <- reactive(AllCounty %>% filter(Mortality >= input$mortality))
    output$plot = renderPlot({
    ggplot(data = AllCounty, aes_string(x = 'long', y = 'lat', group = 'group', fill = input$covariate)) + 
   scale_color_discrete(name = input$covariate) +
                geom_polygon(color = "black") +
               scale_fill_gradientn(colors = brewer.pal(9, "Reds"), trans = ifelse(input$covariate %in% c('Income', 'House_Value', 'Population_Density'),
                 "log10",
                 "identity")) +
   theme(panel.grid.major = element_blank(), 
         panel.background = element_blank(),
         axis.title = element_blank(), 
         axis.text = element_blank(),
         axis.ticks = element_blank())+
          coord_fixed(1.3) +
        ggtitle(paste(
          ifelse(input$covariate %in% c('PM2.5',
                      'Winter_Temperature',
                      'Summer_Temperature',
                      'Population_Density' ,
                      'Income', 
                      "House_Value"), "Average", "Proportion"), 
          input$covariate, ifelse(input$covariate %in% c('Income', 'House_Value', 'Population_Density'), "(log scale)", "")))
    })
    
    output$plot_static = renderPlot({
    ggplot(data = AllCounty, aes_string(x = 'long', y = 'lat', group = 'group', fill = 'Mortality')) + 
   scale_color_discrete(name = "Mortality") +
                geom_polygon(color = "black") +
                scale_fill_gradientn(colors = brewer.pal(9, "Reds")) +
   theme(panel.grid.major = element_blank(), 
         panel.background = element_blank(),
         axis.title = element_blank(), 
         axis.text = element_blank(),
         axis.ticks = element_blank())+
          coord_fixed(1.3) +
        ggtitle(paste("Covid-19 Deaths per 100,000"))
    })

    
        output$slider = renderPlot({
         ggplot(data = AllCounty, aes(x = long, y = lat, group = group)) +
              geom_polygon(color = "red", fill = NA) +
        
              geom_polygon(data= selected(), aes_string(x = 'long', y = 'lat', group = 'group', fill = 'Mortality'))+
		      scale_color_discrete(name = "Mortality") +
                scale_fill_gradientn(colors = brewer.pal(9, "Blues"), trans = "log10") +
   theme(panel.grid.major = element_blank(), 
         panel.background = element_blank(),
         axis.title = element_blank(), 
         axis.text = element_blank(),
         axis.ticks = element_blank())+
          coord_fixed(1.3) +
       ggtitle(paste("Counties with at Least", input$mortality, "Reported Deaths per 100,000 by 10/31/21"))+
            ylab("Deaths per 100,000") +
            xlab(NULL)
          
    })
    
    
}

)

```



