---
title: "Web Scraping"
output: html_document

# Data Wrangling and Web Scraping Code for BST260 Final Project
# Code was modified based on open source provided by: https://github.com/wxwx1993/PM_COVID
# Previous study have been published as "Wu, X., Nethery, R.C., Sabath, M.B., Braun, D. and Dominici, F., 2020. Air pollution and COVID-19 mortality in the United States: strengths and limitations of an ecological regression analysis. Science advances, 6(45), p.eabd4049"
# We have contacted the authors and got permission to use their datasets.
# We thank all of the authors for making their data public and for enabling our project to be possible.
---

```{r}
library(dplyr)
library(stringr)
library(RCurl)
library(httr)

date_of_study <- "10-31-2021"

#Use Apr 15 as the Vax cutoff point
#date_of_study <- "04-15-2021"

```

## Import and Prepare Datasets

We obtained COVID-19 death counts for each county in the United States from Johns Hopkins University, Center for Systems Science and Engineering Coronavirus Resource Center. We collected county-level COVID-19 data reported by the CDC and state health departments, including the number of new and cumulative deaths and confirmed cases reported in each county across the United States, updated daily.

```{r}
# Historical data
covid_historical <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-30-2020.csv"))
covid_us_historical <- subset(covid_historical, Country_Region == "US" & is.na(FIPS) == F)  #Pulled the historical data... 

```

```{r}
# Import COVID Mortality data from JHU CSSE

  df <- read.csv(text = getURL(
        paste0("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/", 
            date_of_study, ".csv")))  #Pulled the raw data up through the 10/31/21
      

    covid <- subset(df, Country_Region == "US")[, 1:12]
    covid <- rbind(covid, subset(covid_us_historical, (!(FIPS %in% covid$FIPS)) & Confirmed == 0 & Deaths == 0 & is.na(FIPS) == F))
    covid$FIPS <- str_pad(covid$FIPS, 5, pad = "0") #...

# Import Long-term Exposure PM2.5 data
county_pm <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_pm25.csv"))
county_temp <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/temp_seasonal_county.csv"))

```

**Import relevant covarites:** We collected various socioeconomic and demographic variables from the 2000 and 2010 Census, the 2005–2016 American Community Surveys and the 2009–2016 CDC Compressed Mortality File.

The American Community Survey data were extracted for each zip code and then averaged across zip codes within a county. We also collect two county-level health risk factors: proportion of residents with obesity and proportion of residents that are current smokers from the Robert Wood Johnson Foundation’s 2020 County Health Rankings. We chose these factors because they would likely predict COVID mortality through biological pathways that are similar to those affected by long-term exposure to PM2.5-- mainly through cardiovascular mechanisms. 

```{r}
# Import census, brfss data as potential confounders
county_census <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/census_county_interpolated.csv"))

GET("https://www.countyhealthrankings.org/sites/default/files/media/document/analytic_data2020.csv", write_disk("county_brfss.csv", overwrite = TRUE))
county_brfss <- read.csv("county_brfss.csv", skip = 1)
county_brfss <- county_brfss[, c('fipscode', 'v011_rawvalue', 'v009_rawvalue')]
names(county_brfss) <- c('fips', 'obese', 'smoke')
county_brfss$fips <- str_pad(county_brfss$fips, 5, pad = "0")

state_test <- read.csv(text = getURL("https://api.covidtracking.com/v1/states/daily.csv"))
state_test <- subset(state_test, date == paste0(substring(str_remove_all(date_of_study, "-"), 5, 8),substring(str_remove_all(date_of_study, "-"), 1, 4)))[, - 38]
statecode <- read.csv(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/statecode.csv"))

hospitals <- read.csv(text = getURL("https://opendata.arcgis.com/datasets/6ac5e325468c4cb9b905f1728d6fbf0f_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D"))
hospitals$BEDS[hospitals$BEDS < 0] <- NA

county_base_mortality <- read.table(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_base_mortality.txt"), sep = "", header = TRUE)
county_old_mortality <- read.table(text = getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_old_mortality.txt"), sep = "", header = TRUE)
county_014_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_014_mortality.txt", sep = "", header = TRUE)
county_1544_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_1544_mortality.txt", sep = "", header = TRUE)
county_4564_mortality <- read.table("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/county_4564_mortality.txt", sep = "", header = TRUE)

colnames(county_old_mortality)[4] <- c("older_Population")
colnames(county_014_mortality)[4] <- c("014_Population")
colnames(county_1544_mortality)[4] <- c("1544_Population")
colnames(county_4564_mortality)[4] <- c("4564_Population")

county_base_mortality <- merge(county_base_mortality,county_old_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_014_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_1544_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)
county_base_mortality <- merge(county_base_mortality,county_4564_mortality[, c(2, 4)], by = "County.Code", all.x = TRUE)

county_base_mortality$older_pecent <- county_base_mortality$older_Population / county_base_mortality$Population
county_base_mortality$"young_pecent" <- county_base_mortality$"014_Population" / county_base_mortality$Population
county_base_mortality$"prime_pecent" <- county_base_mortality$"1544_Population" / county_base_mortality$Population
county_base_mortality$"mid_pecent" <- county_base_mortality$"4564_Population" / county_base_mortality$Population
county_base_mortality$"older_pecent"[is.na(county_base_mortality$"older_pecent")] <- 0
county_base_mortality$"prime_pecent"[is.na(county_base_mortality$"prime_pecent")] <- 0
county_base_mortality$"mid_pecent"[is.na(county_base_mortality$"mid_pecent")] <- 0
county_base_mortality$"young_pecent"[is.na(county_base_mortality$"young_pecent")] <- 0

# Import NCHS Urban-Rural Classification Scheme for Counties
NCHSURCodes2013 <- read.csv("https://raw.githubusercontent.com/wxwx1993/PM_COVID/updated_data/Data/NCHSURCodes2013.csv")
NCHSURCodes2013$FIPS <- str_pad(NCHSURCodes2013$FIPS, 5, pad = "0")

```

## Key Variable Aggregations, Dataset Merge & Join

```{r}
# pm2.5 average over 17 years
county_pm_aggregated <- county_pm %>% 
    group_by(fips) %>% 
    summarise(mean_pm25 = mean(pm25))

# temperature and relative humidity average over 17 years
county_temp_aggregated <- county_temp %>% 
  group_by(fips) %>% 
  summarise(mean_winter_temp = mean(winter_tmmx),
            mean_summer_temp = mean(summer_tmmx),
            mean_winter_rm = mean(winter_rmax),
            mean_summer_rm = mean(summer_rmax))

county_pm_aggregated <- merge(county_pm_aggregated,
                              county_temp_aggregated,
                              by = "fips",
                              all.x = TRUE)

county_hospitals_aggregated <- hospitals %>% group_by(COUNTYFIPS) %>% summarise(beds = sum(BEDS, na.rm = TRUE))
county_hospitals_aggregated$COUNTYFIPS <- str_pad(county_hospitals_aggregated$COUNTYFIPS, 5, pad = "0")

county_census_aggregated2 <- subset(county_census, year == 2016)

# merge dataset
county_census_aggregated2$fips <- str_pad(county_census_aggregated2$fips, 5, pad = "0")
county_census_aggregated2 <- merge(county_census_aggregated2,county_brfss,
                                   by = "fips",
                                   all.x = TRUE)

county_pm_aggregated$fips <- str_pad(county_pm_aggregated$fips, 5, pad = "0")
aggregate_pm <- merge(county_pm_aggregated,covid,
                      by.x = "fips",
                      by.y = "FIPS")

aggregate_pm_census <- merge(aggregate_pm,
                             county_census_aggregated2,
                             by.x = "fips",
                             by.y = "fips")

county_base_mortality$County.Code <- str_pad(county_base_mortality$County.Code, 5, pad = "0")
aggregate_pm_census_cdc <- merge(aggregate_pm_census,
                                 county_base_mortality[, c("County.Code", 
                                                           "Population",
                                                           "older_pecent",
                                                           "young_pecent",
                                                           "prime_pecent",
                                                           "mid_pecent")],
                                 by.x = "fips",
                                 by.y = "County.Code",
                                 all.x = TRUE)

aggregate_pm_census_cdc <- aggregate_pm_census_cdc[is.na(aggregate_pm_census_cdc$fips) == F, ]

#Export Data
write.csv(aggregate_pm_census_cdc,"aggregate_pm_census_cdc_211031.csv", row.names = FALSE)
#write.csv(aggregate_pm_census_cdc,"aggregate_pm_census_cdc_210415.csv", row.names = FALSE)

```





